How many researchers use 'boilerplate' statistical analysis sections?

### 2021-03-15 manuscript/draft_manuscript.Rmd
Update draft_manuscript.Rmd

Two statisticians on this paper (AB and NW) have heard researchers admit that they have copied-and-pasted their statistical methods sections from other papers, regardless of whether they are appropriate. 
The aim of this paper is to use text-mining methods to estimate the extent that researchers are using cut-and-paste or 'boilerplate' statistical methods sections.
Use of these methods sections indicates that little thought has gone into the statistical analysis.

@@

Expanded descriptions of between-group analysis were found in three topics (4, 6 and 10), varying based on method(s) used and study design (Box 3 - include?). Sections assigned to these topics appeared to follow the general template of reporting descriptive statistics followed by methods of hypothesis testing depending on the number of groups being compared. Sections strongly associated with Topic 6 (n = 4,746) combined descriptive statistics with follow-up testing of group differences by Student's t-test, Mann Whitney U or one-way analysis of variance. Descriptive statistics were commonly reported as means with standard deviations or standard errors. Topic 10 (n = 4,664) reported similar methods applied to the analysis of replicates obtained under experimental designs. Variations to descriptions of one-way analysis of variance were identified under Topic 4 (n = 10,163), with an emphasis on post-hoc multiple comparisons.

Remaining topics reflected applications of meta-analysis (Topic 7: n = 5,165), regression modelling (Topic 2: n = 19,478) and larger sections that reported data processing and statistical methods within the same section (Topic 8: n = 38,183). Top matches for Topic 2 shared features of the 'describe-compare' template seen in other topics, expanded to include details of multivariable regression modelling. Examples of multivariable models were (stepwise) logistic regression and cox proportional hazards regression.

@@

We downloaded 28,008 studies. The numbers of excluded studies are shown in Figure~X. 
Of the 12,700 included studies, 9,523 had a statistic section which is 75% (95% CI 74% to 76%).

@@

Variable & Odds ratio & 95% CI
Study type = Observational & 0.78 & 0.69, 0.89
Date (per year) & 0.90 & 0.88, 0.91
Number of funders & 0.80 & 0.74, 0.86
Target sample size (per doubling) & 0.90 & 0.88, 0.92

@@

>>>>

Two statisticians on this paper (AB and NW) have heard researchers admit that they have copied-and-pasted their statistical methods sections from other papers, regardless of whether they are appropriate. 
The aim of this paper is to use text-mining methods to estimate the extent that researchers are using cut-and-paste or 'boilerplate' statistical methods sections.
Boilerplate text is that "which can be reused in new contexts or applications without significant changes to the original" [wikipedia].
Use of these methods sections indicates that little thought has gone into the statistical analysis.

@@

Expanded descriptions of between-group analysis were found in three topics (4, 6 and 10), varying based on method(s) used and study design (Box 3 - include?). Sections assigned to these topics followed the general template of reporting descriptive statistics followed by methods of hypothesis testing depending on the number of groups being compared. Sections strongly associated with Topic 6 (n = 4,746) combined descriptive statistics with follow-up testing of group differences by Student's t-test, Mann Whitney U or one-way analysis of variance. Descriptive statistics were commonly reported as means with standard deviations or standard errors. Topic 10 (n = 4,664) reported similar methods applied to the analysis of replicates obtained under experimental designs. Variations to descriptions of one-way analysis of variance were identified under Topic 4 (n = 10,163), with an emphasis on post-hoc multiple comparisons.

Remaining topics reflected applications of meta-analysis (Topic 7: n = 5,165), regression modelling (Topic 2: n = 19,478) and larger sections that reported data processing and statistical methods within the same section (Topic 8: n = 38,183). Top matches for Topic 2 shared features of the 'describe-compare' template seen in other topics, expanded to include details of multivariable regression modelling. Examples of multivariable models were (stepwise) logistic regression and Cox proportional hazards regression.

@@

We downloaded 28,008 studies. The numbers of excluded studies are shown in Figure~X. 
Of the 12,700 included studies, 9,523 had a statistical methods section, 75% (95% CI 74% to 76%).

@@

```{r}
table =  read.table(header=TRUE, sep='&', text='
Variable & Odds ratio & 95% CI
Study type = Observational & 0.78 & 0.69, 0.89
Date (per year) & 0.90 & 0.88, 0.91
Number of funders & 0.80 & 0.74, 0.86
Target sample size (per doubling) & 0.90 & 0.88, 0.92')
ftab = flextable(table) %>%
theme_box()
autofit()
ftab
```

@@

The clustering algorithm found groups that were purely sample size calculations (topic 2), pilot studies (topic 5) and repeated measures ANOVA (topic 10). There were cases where the exact same method section had been ...
